---
title: "P8105 Homework 3: Vizualizaiton and EDA"
author: "Catherine Lucey (UNI: cml2270)"
date: "10/13/2021"
output: github_document
---

## Setup
Only done once, the setup code loads the necessary libraries and double checks that the working directory is correct. It also sets defaults for figure size with knitr, as well as a default theme and color scale for all ggplots.

```{r setup}
library(tidyverse)
library(readxl)
library(lubridate)
library(ggridges)
getwd()

knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1: Instacart

The Instacart dataset gives the details of many different Instacart orders. There are in total 1,384,617 observations, and 15 variables. Each observation is an item ordered through instacart, and that item is described by the customer order it belongs to, the aisle it came from, the department the aisle is in, the date and time the order was placed, as well as technical information like the customer order ID number and the product ID number. 

```{r load instacart data and plot}

library(p8105.datasets)
data("instacart")

instacart %>% 
  group_by(aisle_id, aisle) %>% 
  summarise(n_obs = n())
```

After grouping by aisle ID number, the 134 rows in the grouped dataframe correspond to the 134 different aisles in the store.

### Aisles Most Commonly Ordered From

Since each item ordered has an associated aisle, the frequency of each aisle value in the "aisle_id" column gives the number of items ordered from each aisle. The aisle data are grouped by department. This plot shows that items are most often ordered from fresh vegetables and fruits aisles, and that the "produce" department overall is the most popular in Instacart orders. The next most popular department is the dairy department, within which the greatest number of items are ordered from the yogurt aise. 

```{r fig.width=10}
instacart %>% 
  group_by(aisle_id, aisle, department_id, department) %>% 
  summarise(n_pdts = n()) %>% 
  filter(n_pdts >= 10000) %>% 
  ggplot(aes(x = reorder(aisle_id, department_id), y = n_pdts, fill = department)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  geom_col(width = 1) +
  geom_text(aes(label = aisle), check_overlap = TRUE, size = 2) +
  theme(legend.position = "left")

```

### Most-Ordered Items in Particular Aisles

The next table shows the three most popular items (as ranked by order frequency) in the dog food and care aisle, the baking ingredients aise, and the packaged vegetable and fruits aise. In the dog food and care aisle, Organix Grain Free Chicken & Vegetable Dog Food	is the most commonly ordered item, light brown sugar is the most popular item in the baking aisle, and organic baby spinach is the most popular item in the packaged fruits and vegetable aisle. The second and third most popular items in each aisle are similar to the first most pouplar item; for example, the third most popular item in the baking and ingredients aisle is granulated white sugar.

```{r}

aisle_popularity_df = 
  instacart %>% 
  filter(aisle == c("packaged vegetables fruits" , "dog food care" , "baking ingredients")) %>% 
  group_by(aisle, product_name) %>% 
  summarise(
    n_ordered = n()) %>% 
  mutate(rank = min_rank(desc(n_ordered))) %>% 
  filter(rank <= 3)

knitr::kable(aisle_popularity_df)

```

### Ice Cream and Apple Orders at Different Times of Day

Looking at when Pink Lady apples and coffee ice cream are typically ordered during the day, coffee ice cream tends to be ordered in the afternoon or evening, wheras pink lady apples are typically ordered in the late morning or early afternoon. Coffee ice cream tends to be ordered earlier in the day on weekends.
    
```{r time of day}

instacart %>% 
  mutate(order_dow = as.character(order_dow)) %>% 
  mutate(order_dow = recode(order_dow, '0' = "Sunday", 
                            '1' = "Monday", 
                            '2' = "Tuesday",
                            '3' = "Wednesday",
                            '4' = "Thursday",
                            '5' = "Friday",
                            '6' = "Saturday")) %>% 
  filter(product_name == c("Pink Lady Apples" , "Coffee Ice Cream")) %>% 
  group_by(order_dow, product_name) %>% 
  summarise(mean_hour = mean(order_hour_of_day)) %>%
  select(order_dow, mean_hour, product_name) %>% 
  pivot_wider(names_from = order_dow, values_from = mean_hour) %>% 
  knitr::kable()
  
```


# Problem 2: BRFSS Data

### Load and Clean the BRFSS data

Tidying the BRFSS data.

```{r load BRFSS data and clean}

library(p8105.datasets)
data("brfss_smart2010")

brfss_overall_df =
  brfss_smart2010 %>% 
  rename(county = Locationdesc, state = Locationabbr) %>%
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  mutate(response = as_factor(response), levels(c("Excellent", "Very Good", "Good", "Fair", "Poor")))

```

### States with 7 or More Study Sites in 2002 and 2010

The table below lists the states that had more than 7 BRFSS study sites in the years 2002 and 2010, as well as lists the number of county-level sites per state in each year. Most states had more than 7 study sites, and in 2010 there were more states with 7+ study sites than in 2002, suggesting that sites were continually added in the study.

```{r}

seven_sites_2002 = 
  brfss_overall_df %>% 
  filter(year == 2002) %>% 
  group_by(state) %>% 
  summarise(n_sites_2002 = n()) %>% 
  filter(n_sites_2002 >= 7)

seven_sites_2010 = 
  brfss_overall_df %>% 
  filter(year == 2010) %>% 
  group_by(state) %>% 
  summarise(n_sites_2010 = n()) %>% 
  filter(n_sites_2010 >= 7)

seven_sites_join_df =
  full_join(seven_sites_2002, seven_sites_2010, by = "state") %>% 
  arrange(state)

knitr::kable(seven_sites_join_df)

```


### Prevalence of "Excellent" Overall Health Responses in BRFSS Data


The following code creates a dataframe giving the crude prevalence of "excellent" overall health ratings in each state (averaged across all state study sites) annually from 2002 to 2010. The plot shows the change in the prevalence of "Excellent" health ratings across time. In this plot, Wyoming study sites had the overall lowest percentage of "excellent" health ratings, and the prevalence of "excellent" responses in Wyoming was also the among the most variable between 2002 and 2010.

```{r}

excellence_df = 
  brfss_overall_df %>%  
  filter(response == "Excellent") %>% 
  group_by(state, year) %>% 
  mutate(avg_val = mean(data_value)) %>% 
  select(response, year, state, avg_val)

excellence_df %>% 
  ggplot(aes(x = year, y = avg_val, group = state, color = state)) + 
  geom_line() +
  labs(y = "Avg Crude Prevalence Per State (Across all Study Sites)",
       x = "Year",
       title = "Percent Prevalence of 'Excellent' Health Status by State Over Time") +
  theme(legend.position = "right")
  
```

### Health Rating Response Values in New York State in 2006 and 2010

In New York State, in both 2006 and 2010, most counties reported a low prevalence of "Poor" overall health ratings among respondents, but the percent prevalence of "Poor" responses was slightly increased in 2010 compared to 2006. "Very good" overall health ratings were the most prevalent responses in both 2006 and 2010, but more counties reported a % prevalence of "Very good" between 30% and 40% in 2002 than in 2010. In both 2006 and 2010, "Excellent" health ratings were moderately prevalent; "excellent" health responses were more prevalent than "poor" and "fair" responses, but less prevalent than "good" and "very good" health ratings.

```{r NYS ratings 2006 2010}

nys_ratings_df =
  brfss_overall_df %>% 
  filter(state == "NY",
         year == "2006" | year == "2010") %>% 
  group_by(county, response)

nys_ratings_df %>% 
  ggplot(aes(x = data_value, fill = response)) +
  geom_density(alpha = 0.4) +
  theme(legend.position = "right") +
  labs(x = "Prevalence of Response (%)") +
  facet_grid(~ year) +
  labs(y = "Proportion of NYS Counties", title = "Percent Prevalence of Health Status Ratings Across NYS Counties in 2006, 2010")
```


# Problem 3

### Tidying and Describing the Acceleromter Data

The accelerometer dataset gives daily minute-by-minute activity counts for five weeks. Thus, in total, there are 35 days' worth of activity counts for every minute of the day; there are 50,400 obvservations total corresponding to the 50,400 minutes in 5 weeks. The week number, day number, and type of day (ex. Friday vs. Wednesday) is given for each minute of observation in addition to the activity count.

```{r load accelerometer data}

accel_df = read_csv("accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_prefix = "activity_",
    names_to = "minute", 
    values_to = "amt_activity") %>%
  group_by(week, day) %>% 
  mutate(day_type = ifelse(day == "Saturday" | day == "Sunday", "weekend", "weekday")) %>% 
  mutate(minute = as.numeric(minute), amt_activity = as.numeric(amt_activity))


```

### Total Activity on Weekdays vs. Weekends

```{r}

daily_totals_df =
  accel_df %>% 
  group_by(day_id, day_type, day) %>% 
  summarise(
    total_activity = sum(amt_activity)
  ) %>% 
  select(day_id, day_type, day, total_activity)
 
daily_totals_df %>% 
  knitr::kable()
  
daily_totals_df %>% 
  ggplot(aes(x = day_id, y = total_activity,fill = day_type)) +
  geom_point() +
  geom_smooth() +
  theme(legend.position = "right")

```

After finding the total activity amount for each day, it can be seen that the maximum amount of activity was `r max(pull(daily_totals_df, "total_activity"))`, which occurred on day 16, a weekday. Plotting the total activity per day and grouping by day type (weekend vs. weekday) shows the participant was generally more active on weekends at the beginning of the study, but by the end of the study at day 35, the participant was more active on weekdays and less active on weekends.

### Minute-by-minute Activity Over the Week

Plotting the minute-by-minute activity counts over the course of 24 hours shows a wide range in the amount of activity during the day. Activity peaks on Friday at about 8 o'clock, and falls again around midnight, when presumably the participant went to bed. It appears that the participant wakes up at around 7AM, as there is a peak of activity at 420 minutes for most days. The participant was also active on Saturday afternoons, when hopefully they were doing something fun.

```{r}

accel_df %>% 
ggplot(aes(x = minute, y = amt_activity, color = day)) +
  geom_point(alpha = 0.5) +
  scale_x_continuous(
    breaks = seq(0, 1440, 60)
  ) +
  geom_smooth() +
  theme(legend.position = "right") +
  labs(x = "Minute (tick marks are hour)", y = "Amount Activity" , title = "Minute-by-Minute Activity Counts Over 24 Hours Per Day Type")
  

```