---
title: "P8105 Homework 3: Vizualizaiton and EDA"
author: "Catherine Lucey (UNI: cml2270)"
date: "10/13/2021"
output: github_document
---

YOU HAVE TO DO THIS: create a subdirectory to store the local data files used in Problems 1 and 2, and use relative paths to access these data files


## Setup
Only done once, the setup code loads the necessary libraries and double checks that the working directory is correct. It also sets defaults for figure size with knitr, as well as a default theme and color scale for all ggplots.

```{r setup}
library(tidyverse)
library(readxl)
library(lubridate)
library(ggridges)
getwd()

knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```
# Problem 1: Instacart

```{r load instacart data and plot}

library(p8105.datasets)
data("instacart")

instacart %>% 
  group_by(aisle_id, aisle) %>% 
  summarise(n_obs = n())
```

After grouping by aisle ID number, the 134 rows in the grouped dataframe correspond to the 134 different aisles in the store.

NOTES: FIX
group by product id ---> number of products --> n_obs should then correspond to the products

```{r fig.width=10}
instacart %>% 
  group_by(aisle_id, aisle, department_id, department) %>% 
  summarise(n_pdts = n()) %>% 
  filter(n_pdts >= 10000) %>% 
  ggplot(aes(x = reorder(aisle_id, department_id), y = n_pdts, fill = department)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  geom_col(width = 1) +
  geom_text(aes(label = aisle), check_overlap = TRUE, size = 2) +
  theme(legend.position = "left")

```

FILTER BY MULTIPLE THINGS USING PIPE |

    Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
    Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).



```{r eval = FALSE}
instacart %>% 
  filter(aisle == "baking ingredients" | "dog food care" | "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarise(
    n_ordered = n(),
    most_ordered = max(product_name)
  )
```


# Problem 2

First, do some data cleaning:

    format the data to use appropriate variable names;
    focus on the “Overall Health” topic
    include only responses from “Excellent” to “Poor”
    organize responses as a factor taking levels ordered from “Poor” to “Excellent”

Using this dataset, do or answer the following (commenting on the results of each):

    In 2002, which states were observed at 7 or more locations? What about in 2010?
    Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
    Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

### Load and Clean the BRFSS data

Tidying the BRFSS data.

```{r load BRFSS data and clean}

library(p8105.datasets)
data("brfss_smart2010")

brfss_overall_df =
  brfss_smart2010 %>% 
  rename(county = Locationdesc, state = Locationabbr) %>%
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  mutate(response = as_factor(response), levels(c("Excellent", "Very Good", "Good", "Fair", "Poor")))

```

Using this dataset, do or answer the following (commenting on the results of each):

    
### States with 7 or More Study Sites in 2002 and 2010

The table below lists the states that had more than 7 BRFSS study sites in the years 2002 and 2010, as well as lists the number of county-level sites per state in each year. Most states had more than 7 study sites, and in 2010 there were more states with 7+ study sites than in 2002, suggesting that sites were continually added in the study.

```{r}

seven_sites_2002 = 
  brfss_overall_df %>% 
  filter(year == 2002) %>% 
  group_by(state) %>% 
  summarise(n_sites_2002 = n()) %>% 
  filter(n_sites_2002 >= 7)

seven_sites_2010 = 
  brfss_overall_df %>% 
  filter(year == 2010) %>% 
  group_by(state) %>% 
  summarise(n_sites_2010 = n()) %>% 
  filter(n_sites_2010 >= 7)

seven_sites_join_df =
  full_join(seven_sites_2002, seven_sites_2010, by = "state") %>% 
  arrange(state)

knitr::kable(seven_sites_join_df)

```


### Prevalence of "Excellent" Overall Health Responses in BRFSS Data

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

The following code creates a dataframe giving the crude prevalence of "excellent" overall health ratings in each state (averaged across all state study sites) annually from 2002 to 2010. The plot shows the change in the prevalence of "Excellent" health ratings across time. In this plot, Wyoming study sites had the overall lowest percentage of "excellent" health ratings, and the prevalence of "excellent" responses in Wyoming was also the among the most variable betwen 2002 and 2010.

```{r}

excellence_df = 
  brfss_overall_df %>%  
  filter(response == "Excellent") %>% 
  group_by(state, year) %>% 
  mutate(avg_val = mean(data_value)) %>% 
  select(response, year, state, avg_val)

excellence_df %>% 
  ggplot(aes(x = year, y = avg_val, group = state, color = state)) + 
  geom_line() +
  labs(y = "Avg Crude Prevalence Per State (Across all Study Sites)",
       x = "Year",
       title = "Percent Prevalence of 'Excellent' Health Status by State Over Time") +
  theme(legend.position = "right")
  
```

### Health Rating Response Values in New York State in 2006 and 2010

In New York State, in both 2006 and 2010, most counties reported a low prevalence of "Poor" overall health ratings among respondents, but the percent prevalence of "Poor" responses was slightly increased in 2010 compared to 2006. "Very good" overall health ratings were the most prevalent responses in both 2006 and 2010, but more counties reported a % prevalence of "Very good" between 30% and 40% in 2002 than in 2010. In both 2006 and 2010, "Excellent" health ratings were moderately prevalent; "excellent" health responses were more prevalent than "poor" and "fair" responses, but less prevalent than "good" and "very good" health ratings.

```{r NYS ratings 2006 2010}

nys_ratings_df =
  brfss_overall_df %>% 
  filter(state == "NY",
         year == "2006" | year == "2010") %>% 
  group_by(county, response)

nys_ratings_df %>% 
  ggplot(aes(x = data_value, fill = response)) +
  geom_density(alpha = 0.5) +
  theme(legend.position = "right") +
  labs(x = "Prevalence of Response (%)") +
  facet_grid(~ year) +
  labs(y = "Proportion of NYS Counties", title = "Percent Prevalence of Health Status Ratings Across NYS Counties in 2006, 2010")
```



# Problem 3


    Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).
    
    Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?
    Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

### Tidying and Describing the Acceleromter Data

The accelerometer dataset gives daily minute-by-minute activity counts for five weeks. Thus, in total, there are 35 days' worth of activity counts for every minute of the day; there are 50,400 obvservations total corresponding to the 50,400 minutes in 5 weeks. The week number, day number, and type of day (ex. Friday vs. Wednesday) is given for each minute of observation in addition to the activity count.

```{r load accelerometer data}

accel_df = read_csv("accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_prefix = "activity_",
    names_to = "minute", 
    values_to = "amt_activity") %>%
  group_by(week, day) %>% 
  mutate(day_type = ifelse(day == "Saturday" | day == "Sunday", "weekend", "weekday")) %>% 
  mutate(minute = as.numeric(minute), amt_activity = as.numeric(amt_activity))


```


```{r}

daily_totals_df =
  accel_df %>% 
  group_by(day_id, day_type, amt_activity) %>% 
  mutate(total_activity = sum(amt_activity))

daily_totals_df %>% 
  ggplot(aes(x = day_id, y = total_activity, color = day_type)) +
  geom_point()

```


i think we could just modify the x axis
```{r}

accel_df %>% 
ggplot(aes(x = minute, y = amt_activity, color = day)) +
         geom_point() 

```


